---
title: "Age of Return"
subtitle: "Markdown for Covariate Data"
author: "Caitlin O'Brien"
date: "1/20/2023"
output:
  rmarkdown::html_document:
    theme: cosmo
    toc: TRUE
    toc_float: FALSE
    code_folding: hide
---


```{r, warning=FALSE, message=FALSE, error=FALSE, class.source='fold-show'}
#Packages
library(tidyverse)
library(here)
library(data.table)
library(zoo)
```

# {.tabset}

## Basefile


**Year Joins used:**

[Shared googlesheet list of joins](https://docs.google.com/spreadsheets/d/1SJ3gYAqOGeecirm4OECOcIXv3z6DvwcpU4EPKHBCpVM/edit#gid=0)

- smolt migration year (smolt) = year of detection at LGR
- adult detection (ADU_first) = return detection at first dam (BON > MCN > ICH > LGR)
- water year (WY) = Oct to Dec of previous calendar year plus Jan to Sept of the current calendar year
- migration year (MY) = used for year prior to entry and year of entry. Includes April to December of previous calendar year plus Jan to March of current calendar year. 

Updated codes:



### Data wrangling:

#### FOR ORIGINAL DATA SET
```{r,warning=FALSE, message=FALSE, error=FALSE}
d<-read_csv(here("data", "data_adults.csv"))


#extract smolt migration year (smolt), smolt_year prior = smolt+1, first adult detection, then separate by month and year
#filter to include data within time frame, size, and rkm

d<-d %>% 
  mutate(pass_type_T_R = dplyr::recode(pass_type_T_R, "TB" = "ROR", "RORBD" = "ROR")) %>% 
  filter(length < 155, #length threshold. Note: leaving in animals less than 55 mm
         juv_lgr_first > '1998-01-01 00:00:00' & juv_lgr_first < '2018-12-31 24:00:00', #set time frame
         rel_rkm >= 522.173) %>% #removes lower tucannon singleton data point--confirm
   mutate(smolt_year = as.character(lubridate::year(juv_lgr_first))) %>% 
  mutate(smolt_entry =as.character(lubridate::year(juv_lgr_first)+1) ) %>% #for entry in new WY
   mutate(adu_first = if_else(!is.na(adu_bon_first) , adu_bon_first , if_else(!is.na(adu_mcn_first),adu_mcn_first,if_else(!is.na(adu_ich_first),adu_ich_first,adu_lgr_first)))) %>%
  mutate(adult_year = as.character(lubridate::year(adu_first))) %>% 
  mutate( smolt_DOY = as.character(lubridate::yday(juv_lgr_first)))
```


#### FOR UPDATED DATA SET
```{r}
d<-read_csv(here("data", "updated_data_adults.csv")) #subset from MC's new data submitted (2.14.23)

d <- d %>% 
  filter(t_rear_type == "W") %>% 
  mutate(smolt_entry =as.character(lubridate::year(juv_lgr_first)+1)) %>%  #for table join: entry in new WY
  mutate(adult_year = as.character(adult_year),
         adult_month = as.character(adult_month), 
         adult_DOY = as.character(adult_DOY),
         smolt_year = as.character(smolt_year), 
         smolt_month = as.character(smolt_month), 
         smolt_DOY = as.character(smolt_DOY)
         ) %>% 
  mutate(BY = as.integer(smolt_year)-1) %>% 
  mutate(BY = as.character(BY))


#paired down file to join covariates
d.cov <- d %>% 
  select("tag_id" , "t_run", "t_rear_type", "length", "smolt_year","smolt_entry", "BY", "adult_year", "smolt_DOY", "locgrp", "years_diff", "OceanAge", "pass_type_T_R")
```


### Visualize
```{r,warning=FALSE, message=FALSE, error=FALSE, fig.show="hold", out.width="50%"}

sum.sm<-d%>% 
  group_by(smolt_month) %>% 
  summarize(sum = n())

plot(sum.sm$smolt_month, sum.sm$sum, type="b")

sum.am<-d %>% 
  group_by(adult_month) %>% 
  summarize(sum = n())

plot(sum.am$adult_month, sum.am$sum, type="b") # Filter by months?
```



## A.	Large-scale Ocean/climate indices 



### **1.	NPGO (North Pacific Gyre Oscillation) index**

*	Mean of ONDJFM of first fall/winter of ocean entry
    + SMY+1:WY --> updated with JG code BY:WY
-	Mean of ONDJFM of last fall/winter in ocean before adult return 
    + ADU:WY

```{r}
NPGO.data <- read_csv(here("Data", "NPGO 1990-2022.csv"))

NPGO<- NPGO.data %>% 
  filter( month %in% c(10:12, 1:3)) %>%
  mutate( wateryear = ifelse(month %in% c(1:9), year, year+1)) %>% 
  group_by(wateryear) %>% 
  summarize( meanONDJFM = round(mean(NPGO),3)) %>% 
  mutate(wateryear = as.character(wateryear))
```


```{r eval=FALSE, include=FALSE}

###original code--use Jenn's, see below
#Mean of ONDJFM of first fall/winter of ocean entry
#Add mean for first year entry to basefile
d.cov<- d.cov%>% 
  inner_join(NPGO, by = c("smolt_entry" = "wateryear")) %>% 
  rename(meanNPGO_ONDJFM_smoltentry = meanONDJFM)

##Mean of ONDJFM of last fall/winter in ocean before adult return 
##Add mean for last fall/winter before adult return to basefile
d.cov<- d.cov %>% 
  inner_join(NPGO, by = c("adult_year" = "wateryear")) %>% 
  rename(meanNPGO_ONDJFM_adu = meanONDJFM)


d.cov %>% 
  select("smolt_year", "smolt_entry", "adult_year","meanNPGO_ONDJFM_smoltentry", "meanNPGO_ONDJFM_adu") %>% 
  head(.,10)
```

```{r}

#NPGO, Ocean conditions setup prior to smolt ocean entry, therefore in BY of spring/summer Chinook smolts=
#Mean of ONDJFM of first fall/winter of ocean entry
d.cov<- d.cov%>% 
  inner_join(NPGO, by = c("smolt_year" = "wateryear")) %>% 
  rename(NPGO.ONDJFM.0 = meanONDJFM)

##Mean of ONDJFM of last fall/winter in ocean before adult return 
d.cov<- d.cov %>% 
  inner_join(NPGO, by = c("adult_year" = "wateryear")) %>% 
  rename(NPGO.ONDJFM.T = meanONDJFM)


d.cov %>% 
  select("smolt_year", "adult_year","NPGO.ONDJFM.0", "NPGO.ONDJFM.T") %>% 
  head(.,10)

```



### **2.	PDO (Pacific Decadal Oscillation) index**

-	Monthly data
-	Mean of AMJJAS of spring/summer in year of ocean entry
    + smolt:WY --> changed to BY:WY
-	Mean of ONDJFM of first fall/winter in ocean --> changed to fall/winter prior to smolt ocean entry
    + smolt+1:WY --> chnaged to smolt : WY
-	Mean of ONDJFM of last fall/winter in ocean before adult return
    + ADU:WY

```{r}
ERSST.v5.PDO <- fread("https://www.ncei.noaa.gov/pub/data/cmb/ersst/v5/index/ersst.v5.pdo.dat")

PDO<- ERSST.v5.PDO %>% 
  pivot_longer(cols = c(2:13),
               names_to = "month",
               values_to = "PDO") %>% 
  mutate(month = match(month,month.abb)) %>% 
  mutate( wateryear = ifelse(month %in% c(1:9), Year, (Year+1))) 

#Mean of ONDJFM of first fall/winter in ocean
PDO.ONDJFM<- PDO%>% 
  filter(month %in% c(1:3,10:12)) %>% 
  group_by(wateryear) %>% 
  summarize( meanPDO= round(mean(PDO),3)) %>% 
  mutate(wateryear = as.character(wateryear))

#Mean of AMJJAS of spring/summer in year of ocean entry
PDO.AMJJAS <- PDO %>% 
  filter(month %in% c(4:9)) %>% 
  group_by(wateryear) %>% 
  summarize( meanPDO= round(mean(PDO),3)) %>% 
  mutate(wateryear = as.character(wateryear))
  
```

old
```{r eval=FALSE, include=FALSE}
#Mean of AMJJAS of spring/summer in year of ocean entry 
d.cov<- d.cov %>% 
  inner_join(PDO.AMJJAS, by = c("smolt_year" = "wateryear")) %>% 
  rename(PDO.AMJJAS.1 = meanPDO)

#Mean of ONDJFM of first fall/winter in ocean

d.cov<- d.cov %>% 
  inner_join(PDO.ONDJFM, by = c("smolt_entry" = "wateryear")) %>% 
  rename(PDO.ONDJFM. = meanPDO)

#Mean of ONDJFM of last fall/winter in ocean before adult return
d.cov<- d.cov %>% 
  inner_join(PDO.ONDJFM, by = c("adult_year" = "wateryear")) %>% 
  rename(meanPDO_ONDJFM_adu = meanPDO)

d.cov %>% 
  select("smolt_year", "smolt_entry","adult_year","meanPDO_AMJJAS_smolt", "meanPDO_ONDJFM_smoltentry", "meanPDO_ONDJFM_adu") %>% 
  head(.,10)

```
updated 
```{r}
#Mean of AMJJAS of spring/summer in year of ocean entry 
d.cov<- d.cov %>% 
  inner_join(PDO.AMJJAS, by = c("BY" = "wateryear")) %>% 
  rename(PDO.AMJJAS.1 = meanPDO)

#Mean of ONDJFM of first fall/winter in ocean //changed to// fall/winter prior to smolt ocean entry

d.cov<- d.cov %>% 
  inner_join(PDO.ONDJFM, by = c("smolt_year" = "wateryear")) %>% 
  rename(PDO.ONDJFM.0 = meanPDO)

#Mean of ONDJFM of last fall/winter in ocean before adult return
d.cov<- d.cov %>% 
  inner_join(PDO.ONDJFM, by = c("adult_year" = "wateryear")) %>% 
  rename(PDO.ONDJFM.T = meanPDO)

d.cov %>% 
  select("smolt_year", "BY","adult_year","PDO.AMJJAS.1", "PDO.ONDJFM.0", "PDO.ONDJFM.T") %>% 
  head(.,10)

```


### **3.	SSTarc data**
-	File: File on local drive until data is released -> ersstArc.csv (see shared dropbox)

Could try similar to PDO covariates:

  - Mean of AMJJAS of spring/summer in year of ocean entry
    + smolt:WY
  - Mean of ONDJFM of first fall/winter in ocean 
    + smolt+1:WY
  - Mean of ONDJFM of last fall/winter in ocean before adult return
    + ADU:WY



```{r, warning=FALSE, message=FALSE, error=FALSE}
SSTarc<-read_csv(here("data", "ersstArc.csv")) # select ersstArc.csv in shared dropbox

SSTarc.AMJJAS<-SSTarc %>% 
  mutate( wateryear = ifelse(month %in% c(1:9), year, (year+1))) %>% 
  filter(month %in% c(4:9)) %>% 
  group_by(wateryear) %>% 
  summarize( meanSSTarc= round(mean(sstarc),3)) %>% 
  mutate(wateryear = as.character(wateryear))

SSTarc.ONDJFM<-SSTarc %>% 
  mutate( wateryear = ifelse(month %in% c(1:9), year, (year+1))) %>% 
  filter(month %in% c(1:3,10:12)) %>% 
  group_by(wateryear) %>% 
  summarize( meanSSTarc= round(mean(sstarc),3)) %>% 
  mutate(wateryear = as.character(wateryear))

```

```{r}

#Mean of AMJJAS of spring/summer in year of ocean entry 
d.cov<- d.cov %>% 
  inner_join(SSTarc.AMJJAS, by = c("smolt_year" = "wateryear")) %>% 
  rename(meanSSTarc_AMJJAS_smolt = meanSSTarc)

#Mean of ONDJFM of first fall/winter in ocean

d.cov<- d.cov %>% 
  inner_join(SSTarc.ONDJFM, by = c("smolt_entry" = "wateryear")) %>% 
  rename(meanSSTarc_ONDJFM_smoltentry = meanSSTarc)

#Mean of ONDJFM of last fall/winter in ocean before adult return
d.cov<- d.cov %>% 
  inner_join(SSTarc.ONDJFM, by = c("adult_year" = "wateryear")) %>% 
  rename(meanSSTarc_ONDJFM_adu = meanSSTarc)

d.cov %>% 
  select("smolt_year", "smolt_entry","adult_year","meanSSTarc_AMJJAS_smolt", "meanSSTarc_ONDJFM_smoltentry", "meanSSTarc_ONDJFM_adu") %>% 
  head(.,10)

```

### **4.	AL (Aleutian Low) index**

-	Download monthly data
-	Mean of ONDJFM of first fall/winter in ocean
    + smolt+1 : WY
-	Mean of ONDJFM of fall/winter prior to ocean entry
    + smolt:WY
-	Mean of ONDJFM of last fall/winter in ocean before adult return
    + ADU:WY
-	More info & to download data https://psl.noaa.gov/data/timeseries/ALBSA/ 


```{r, warning=FALSE, message=FALSE, error=FALSE}

AL<-fread("https://psl.noaa.gov/data/timeseries/monthly/data/albsa.ncepr1.data", header = FALSE )
colnames(AL)<-c("year",paste0(1:12))


AL<-AL %>% 
  pivot_longer(cols = c(2:13),
               names_to = "month",
               values_to = "AL") %>% 
  mutate(month = as.integer(month)) %>% 
  mutate( wateryear = ifelse(month %in% c(1:9), year, (year+1))) %>% 
  filter(month %in% c(1:3,10:12)) %>% 
  group_by(wateryear) %>% 
  summarise(meanAL = mean(AL)) %>% 
  mutate(wateryear = as.character(wateryear))
  

```

```{r}

#Mean of ONDJFM of fall/winter prior to ocean entry 
d.cov<- d.cov%>% 
  inner_join(AL, by = c("smolt_year" = "wateryear")) %>% 
  rename(meanAL_ONDJFM_smolt = "meanAL") 

#Mean of ONDJFM of first fall/winter in ocean
d.cov<- d.cov%>% 
  inner_join(AL, by = c("smolt_entry" = "wateryear")) %>% 
  rename(meanAL_ONDJFM_smoltentry = "meanAL") 


#Mean of ONDJFM of last fall/winter in ocean before adult return
d.cov<- d.cov %>% 
  inner_join(AL, by = c("adult_year" = "wateryear")) %>% 
  rename(meanAL_ONDJFM_adu = "meanAL") 

d.cov %>% 
  select("smolt_year", "smolt_entry","adult_year","meanAL_ONDJFM_smolt", "meanAL_ONDJFM_smoltentry", "meanAL_ONDJFM_adu") %>% 
  head(.,10)
```



### **5.	ENSO 3.4**

-	Download monthly data
-	Mean of year prior to ocean entry
    + ?
    + currently using smolt:MY
-	Mean of MJJAS of year of ocean entry
    + smolt:WY
-	Mean of DJF of first winter in ocean
    + smolt+1:WY
-	Mean of DJF of last winter in ocean before adult return
    + ADU:WY
-	More info & to download data https://climatedataguide.ucar.edu/climate-data/nino-sst-indices-nino-12-3-34-4-oni-and-tni 

```{r, warning=FALSE, message=FALSE, error=FALSE}
ENSO<-fread("http://www.esrl.noaa.gov/psd/data/correlation/nina34.data", header = FALSE )
colnames(ENSO)<-c("year",paste0(1:12))

#summarize by mean ENSO per year
ENSO.year<-ENSO %>% 
  pivot_longer(cols = c(2:13),
               names_to = "month",
               values_to = "ENSO") %>% 
  mutate(month = as.integer(month)) %>% 
  mutate( MY = ifelse(month %in% c(1:3), year, (year+1))) %>% #A:D of current year plus J:M of upcoming year = MY
  group_by(MY) %>% 
  summarise(meanENSO = mean(ENSO)) %>% 
  mutate(MY = as.character(MY))


#summarize by DJF
ENSO.DJF<-ENSO %>% 
  pivot_longer(cols = c(2:13),
               names_to = "month",
               values_to = "ENSO") %>% 
  mutate(month = as.integer(month)) %>% 
  mutate( wateryear = ifelse(month %in% c(1:9), year, (year+1))) %>% 
  filter(month %in% c(1,2,12)) %>% 
  group_by(wateryear) %>% 
  summarise(meanENSO = mean(ENSO)) %>% 
  mutate(wateryear = as.character(wateryear))

#summarize by MJJAS
ENSO.MJJAS<-ENSO %>% 
  pivot_longer(cols = c(2:13),
               names_to = "month",
               values_to = "ENSO") %>% 
  mutate(month = as.integer(month)) %>% 
  mutate( wateryear = ifelse(month %in% c(1:9), year, (year+1))) %>% 
  filter(month %in% c(5:9)) %>% 
  group_by(wateryear) %>% 
  summarise(meanENSO = mean(ENSO)) %>% 
  mutate(wateryear = as.character(wateryear))
```


```{r}
#Mean of year prior to ocean entry (doublecheck, using April to March as 'year prior')
d.cov<- d.cov%>% 
  inner_join(ENSO.year, by = c("smolt_year" = "MY")) %>% 
  rename(meanENSO_smolt = "meanENSO") 

#Mean of MJJAS of year of ocean entry 

d.cov<- d.cov%>% 
  inner_join(ENSO.MJJAS, by = c("smolt_year" = "wateryear")) %>% 
  rename(meanENSO_MJJAS_smolt = "meanENSO") 

#Mean of DJF of first winter in ocean
d.cov<- d.cov%>% 
  inner_join(ENSO.DJF, by = c("smolt_entry" = "wateryear")) %>% 
  rename(meanENSO_DJF_smoltentry = "meanENSO") 

#Mean of DJF of last winter in ocean before adult return
d.cov<- d.cov %>% 
  inner_join(ENSO.DJF, by = c("adult_year" = "wateryear")) %>% 
  rename(meanENSO_DJF_adu = "meanENSO") 

d.cov %>% 
  select("smolt_year", "smolt_entry","adult_year","meanENSO_smolt", "meanENSO_MJJAS_smolt", "meanENSO_DJF_smoltentry", "meanENSO_DJF_adu") %>% 
  head(.,10)
```





## B.	Coastal/local indices

### **1.	CUTI**

-	Annual index
-	Year of ocean entry
    + ? 
    + currently using smolt:MY
-	More info & to download data  https://mjacox.com/upwelling-indices/ 
+ selected 47N as the northmost location of CUTI/BEUTI versus 46N due to SRSC

```{r}
CUTI<-fread("https://www.mjacox.com/wp-content/uploads/2023/01/CUTI_monthly.csv")

CUTI<- CUTI %>% 
  select(1, 2,19) %>% 
  mutate( MY = ifelse(month %in% c(1:3), year, (year+1))) %>%  #using April to March for MY
  rename(CUTI = 3) %>% 
   group_by(MY) %>% 
  summarise(meanCUTI= round(mean(CUTI),3)) %>%  #get mean of 47N location across months per year
  mutate(MY = as.character(MY))
```

```{r}
#Mean of year of ocean entry
d.cov<- d.cov%>% 
  inner_join(CUTI, by = c("smolt_year" = "MY")) %>% 
  rename(meanCUTI_smolt = "meanCUTI") 

d.cov %>% 
  select("smolt_year", "meanCUTI_smolt") %>% 
  head(.,10)
```



### **2.	BEUTI**

-	Annual index
-	Year of ocean entry
    + ? 
    + currently using smolt:MY
-	More info & to download data https://mjacox.com/upwelling-indices/ 

```{r}
BEUTI<-fread("https://www.mjacox.com/wp-content/uploads/2023/01/BEUTI_monthly.csv")

BEUTI<- BEUTI %>% 
  select(1, 2,19) %>% #selecting location 47N only
  mutate( MY = ifelse(month %in% c(1:3), year, (year+1))) %>%  #using April to March for MY
  rename(BEUTI = 3) %>% 
   group_by(MY) %>% 
  summarise(meanBEUTI= round(mean(BEUTI),3)) %>%  #mean of 47N location across months per year
  mutate(MY = as.character(MY))
```

```{r}
#Mean of year of ocean entry
d.cov<- d.cov%>% 
  inner_join(BEUTI, by = c("smolt_year" = "MY")) %>% 
  rename(meanBEUTI_smolt = "meanBEUTI") 

d.cov %>% 
  select("smolt_year", "meanBEUTI_smolt") %>% 
  head(.,10)
```




## C.	Regional freshwater indices

### **1.	PNI (Pacific Northwest Index)**

-	Annual index
- Mean of year of ocean entry
    + ? smolt:CY
-	More info & to retrieve data https://www.cbr.washington.edu/status/pni 
*	Use the DFA-derived index (left-most column); this index may need revising


```{r, warning=FALSE, message=FALSE, error=FALSE}
PNI<- read_csv(here("data", "PNI2021.csv"), skip=1)

PNI<-PNI %>% 
  select(1, 2) %>% 
  rename(year = 1, PNI = 2) %>% 
  mutate(year = as.character(year))
```

```{r}
#Mean of year of ocean entry
d.cov<- d.cov%>% 
  inner_join(PNI, by = c("smolt_year" = "year")) %>% 
  rename(meanPNI_smolt = "PNI") 

d.cov %>% 
  select("smolt_year", "meanPNI_smolt") %>% 
  head(.,10)
```


### **2.	Snow-water-equivalent data from SNOTEL**

Data online

-	Annual index
- Mean of year of ocean entry
    + ? smolt:CY (confirm data was aggregated by CY and not WY)
-	File: swepeak.wnb2021.csv (in Dropbox)
-	Use yearly meanSWE data


```{r, warning=FALSE, message=FALSE, error=FALSE}
SWE<-read_csv(here("data", "swepeak.wnb2021.csv"))

SWE<-SWE %>% 
  select(1, 7) %>% 
  mutate(year = as.character(year))
```

```{r}
#Mean of year of ocean entry
d.cov<- d.cov%>% 
  inner_join(SWE, by = c("smolt_year" = "year")) %>% 
  rename(meanSWE_smolt = "meanSWE") 

d.cov %>% 
  select("smolt_year", "meanSWE_smolt") %>% 
  head(.,10)
d.cov
```






## D.	Local freshwater indices

### **LGR all indices:**
+ smolt, DOY : CY, DOY 

**1.	LGR WQM river temperature**

-	Daily data
-	Calculate a 7-d running mean, right-aligned to LGR passage date (i.e., 7th day of the mean is the day the fish passed LGR)

**2.	LGR flow**

-	Daily data
-	Calculate a 7-d running mean, right-aligned to LGR passage date

**3.	LGR % spill**

-	Daily data
-	Calculate a 7-d running mean, right-aligned to LGR passage date

**4.	LGR % TDG**

-	Daily data
-	Calculate a 7-d running mean, right-aligned to LGR passage date

```{r, warning=FALSE, message=FALSE, error=FALSE}
LGR<-fread("https://www.cbr.washington.edu/dart/cs/php/rpt/mg.php?sc=1&mgconfig=river&outputFormat=csv&year%5B%5D=2022&year%5B%5D=2021&year%5B%5D=2020&year%5B%5D=2019&year%5B%5D=2018&year%5B%5D=2017&year%5B%5D=2016&year%5B%5D=2015&year%5B%5D=2014&year%5B%5D=2013&year%5B%5D=2012&year%5B%5D=2011&year%5B%5D=2010&year%5B%5D=2009&year%5B%5D=2008&year%5B%5D=2007&year%5B%5D=2006&year%5B%5D=2005&year%5B%5D=2004&year%5B%5D=2003&year%5B%5D=2002&year%5B%5D=2001&year%5B%5D=2000&year%5B%5D=1999&year%5B%5D=1998&loc%5B%5D=LWG&data%5B%5D=Dissolved+Gas+Percent&data%5B%5D=Outflow&data%5B%5D=Spill+Percent&data%5B%5D=Temp+%28WQM%29&startdate=1%2F1&enddate=12%2F31&avgyear=0&consolidate=1&grid=1&y1min=0&y1max=&y2min=&y2max=&size=medium")

LGR.rmean<-LGR %>% 
  #mutate(across(everything(), ~ifelse(.=="", NA, as.character(.)))) %>% 
    pivot_longer(cols = c(2:101),
               names_to = "ID",
               values_to = "value") %>% 
  separate(ID, c('year', 'LGR', 'ID')) %>% 
    pivot_wider(names_from = "ID",
             values_from = "value") %>% 
  mutate_at(c("gaspct", "outflow", "spillpct", "tempc"), as.numeric)  %>% 
  separate(1, c('mm',"dd")) %>% 
  mutate(mm = as.numeric (mm),
         year = as.numeric(year), 
         dd = as.numeric(dd)) %>% 
arrange(year) %>% #sort data to use rollmean
  group_by(year) %>% 
  mutate(TDG_7d = rollmean(gaspct, k=7, fill=NA, align = "right"), #check rollmean v rollapply for handling NAs
         Flow_7d = rollmean(outflow, k=7, fill=NA, align = "right"),
            Spill_7d = rollmean(spillpct, k=7, fill=NA, align = "right"),
           Temp_7d = rollmean(tempc, k=7, fill=NA, align = "right")) %>% 
  mutate(YMD = lubridate::ymd(paste(year,mm,dd, sep ="-"))) %>% 
   mutate( DOY = lubridate::yday(YMD)) %>% 
  select("year", "DOY", "TDG_7d", "Flow_7d","Spill_7d", "Temp_7d") %>% 
   mutate(year = as.character(year),
          DOY = as.character(DOY))


#need to add centering and scaling?
```

```{r}

###join by year and doy fish passed LGR
d.cov <- d.cov %>% 
  inner_join(LGR.rmean, by = c("smolt_year" = "year", "smolt_DOY" = "DOY"))  


d.cov %>% 
  select("smolt_year", "smolt_DOY", "TDG_7d", "Flow_7d","Spill_7d", "Temp_7d") %>% 
  head(., 10) #Need to look into NA's---replace with previous average?
```




## E.	Fish covariates: 

(determined from PTAGIS data set)

**1.	LGR passage timing**

 + smolt_DOY

**2.	Release location groupings (an index of Major Population Group?)**

-	Method TBD

- To use HUC6
```{r}
#add HUC8 to missing rel_huc:
df$rel_huc[df$rel_site == "SNAKE3" & is.na(df$rel_huc)]<-"17060103"
#SNAKE3	Snake River - Clearwater River to Salmon River (km 224-303)	17060103	Lower Snake-Asotin	Lower Snake	522; snake river trap @ 522.225
df$rel_huc[df$rel_site == "GRANDR" & is.na(df$rel_huc)]<-"17060106"
#GRNTRP	Grande Ronde River Trap	17060106	Lower Grande Ronde	Lower Snake	522.271.002


#add deliminator
df$split<-gsub("(\\d{2})$",",\\1",df$rel_huc)

#separate
df<-separate(df, split, into = c("HUC6", "HUC8")) 

#rename at basin level
df<-df %>% 
  mutate(basin = case_when(HUC6 == "170601" ~ "Lower Snake", 
                   HUC6 == "170602" ~ "Salmon", 
                   HUC6 == "170603" ~ "Clearwater"))
```


**3.	Above LGR vs At LGR groupings**

- To use at/above LGR see column `locgrp`

**4.	Transported vs in-river passage type**

+ relabelled all to either T or ROR

**5.	Smolt fork length**

*filtered >155 mm, keeping smaller than 55 mm but >25 mm for now. 



## Save file
```{r}
write.csv(d.cov, file=here("data", "covariate_data_age.csv"), row.names = FALSE)
```

